### 2020-05-29 writeup/proof_main.tex
minor

\begin{theorem}\label{thm_model_shift}
For $i=1,2$, let $Y_i = X_i\beta_i + \varepsilon_i$ be two independent data models, where $X_i=Z_i\Sigma_i^{1/2}\in \R^{n_i\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_i:=n_i/p>1$ being fixed constants, and $\e_i\in \R^{n_i}$ are  random vectors with i.i.d. entries with mean zero, variance $\sigma^2$ and all moments as in \eqref{assmAhigh}.  
	Then with high probability, we have
	\begin{align}
	 	\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \ge  \delta \label{upper}\\
		\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \le -\delta, \label{lower}
	\end{align}
	where
	\begin{align} %\bigtr{{\Sigma_2^{-1}}}
		\Delta_{\vari} &\define {\sigma^2}\bigbrace{\frac{1}{\rho_2 - 1} -  \frac{1}{\rho_1 + \rho_2}\cdot \frac1p \bigtr{(\hat a_1 \hat M^{\top}\hat M + \hat a_2\id)^{-1}} }  \\
		\Delta_{\beta} &\define (\beta_s - \hat{v}\beta_t)^{\top} \Sigma_1^{1/2} \Pi \Sigma_1^{1/2} (\beta_s - \hat{v}\beta_t). 
	\end{align}
\end{theorem}

@@

\begin{proof}[Proof of Theorem \ref{thm_many_tasks}]
	By fixing $W_1, W_2, \dots, W_k$, we can derive a closed form solution for $B$ as
	\begin{align*}
		\hat{B}(W_1, \dots, W_k) &= (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{i=1}^k Y_i W_i^{\top}} (Z Z^{\top})^{-1} \\
		&= \sum_{i=1}^k \bigbrace{\beta_i W_i^{\top}} (ZZ^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1}
	\end{align*}
	where we denote $Z\in\real^{r\times k}$ as the $k$ vectors $W_1, W_2, \dots, W_k$ stacked together.
	%Now we switch $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} to
	Similar to Section \ref{sec_setup}, we consider minimizing the validation loss over $W_1, W_2, \dots, W_k$ provided with $\hat{B}$.

>>>>

\begin{theorem}\label{thm_model_shift}
For $i=1,2$, let $Y_i = X_i\beta_i + \varepsilon_i$ be two independent data models, where $X_i=Z_i\Sigma_i^{1/2}\in \R^{n_i\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_i:=n_i/p>1$ being fixed constants, and $\e_i\in \R^{n_i}$ are random vectors with i.i.d. entries with mean zero, variance $\sigma^2$ and all moments as in \eqref{assmAhigh}. Moreover suppose that $X_i$, $\beta_i$ and $\varepsilon_i$ are independent of each other. 
	Then with high probability, we have
	\begin{align}
	 	\te(\hat{\beta}_{t}^{\MTL}) \le \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \ge  \delta \label{upper}\\
		\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL}) \quad \text{ when: } \ \ &\Delta_{\vari} - \Delta_{\beta} \le -\delta, \label{lower}
	\end{align}
	where
	\begin{align} %\bigtr{{\Sigma_2^{-1}}}
		\Delta_{\vari} &\define {\sigma^2}\bigbrace{\frac{1}{\rho_2 - 1} -  \frac{1}{\rho_1 + \rho_2}\cdot \frac1p \bigtr{(\hat a_1 \hat M^{\top}\hat M + \hat a_2\id)^{-1}} }  \\
		\Delta_{\beta} &\define (\beta_s - \hat{v}\beta_t)^{\top} \Sigma_1^{1/2} \Pi \Sigma_1^{1/2} (\beta_s - \hat{v}\beta_t). 
	\end{align}
\end{theorem}

@@

\begin{proof}[Proof of Theorem \ref{thm_many_tasks}]
\todo{check!} By fixing $W_1, W_2, \dots, W_k$, we can derive a closed form solution for $B$ as
	\begin{align*}
		\hat{B}(W_1, \dots, W_k) &= (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{i=1}^k Y_i W_i^{\top}} (Z Z^{\top})^{-1} \\
		&= \sum_{i=1}^k \bigbrace{\beta_i W_i^{\top}} (ZZ^{\top})^{-1} + (X^{\top}X)^{-1}X^{\top} \bigbrace{\sum_{i=1}^k \varepsilon_i W_i^{\top}} (ZZ^{\top})^{-1}
	\end{align*}
	where we denote $Z\in\real^{r\times k}$ as the $k$ vectors $W_1, W_2, \dots, W_k$ stacked together.
	%Now we switch $\hat{B}$ back into equation \eqref{eq_mtl_same_cov} to
	Similar to Section \ref{sec_setup}, we consider minimizing the validation loss over $W_1, W_2, \dots, W_k$ provided with $\hat{B}$.

### 2020-05-29 writeup/proof_taskonomy.tex
minor

\section{Supplementary Materials for Section \ref{sec_taskonomy}}\label{app_proof_sec4}

\begin{proposition}\label{prop_taskonomy}
Let $X_i \in\real^{n_i \times p}$ and $Y_i = X_i\beta_i + \varepsilon_i$, for $i = 1, \cdots, t$, where $\e_i$ are random vectors with i.i.d. entries of mean zero and variance $\sigma^2$. Let $\rho_i>1$, $1\le i \le t$ be fixed constants. Assume that $\max_{1\le i \le t}\|\beta_i\|=\OO(1)$, $\sigma^2=\OO(1)$ and 
\be\label{BTBassm}
\left\|\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1}\right\|=\OO(1),\quad B_{t-1}^{\star}: = [{\beta}_1,{\beta}_2,\dots,{\beta}_{t-1}].
\ee
Suppose that all random variables have subexponential decay. Then we have with high probability,
\be\label{eq_taskonomy}
\te(B W_t) =\|\e_\beta + \sigma^2 \e_{\vari}^{(1)}\|^2+ \sigma^2 \|\e_{\vari}^{(2)}\|^2 + \OO(p^{-1/2+\e}),
\ee
where $B^{\star}_{t-1} := [{\beta}_1,{\beta}_2,\dots,{\beta}_{t-1}]$, and
\begin{align*}
&\e_\beta:= \Sigma_k^{1/2}\left\{1 - B^\star_{t-1}\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1} (B^\star_{t-1})^\top\right\} \beta_t,\\
&\e_{\vari}^{(1)}:=  \Sigma_k^{1/2}  B^\star_{t-1}\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1}\cal M_1 \left[(B^\star_{t-1})^\top B^\star_{t-1} + \sigma^2\cal M_1\right]^{-1} (B^\star_{t-1})^\top \beta_t,\\
&\e_{\vari}^{(2)}:= \cal M_2^{1/2}\left[(B^\star_{t-1})^\top B^\star_{t-1} + \sigma^2\cal M_1\right]^{-1} (B^\star_{t-1})^\top \beta_t.
\end{align*}
Here $\cal M_1$ and $\cal M_2$ are $k\times k$ diagonal matrices with $(\cal M_1)_{ii}=\frac{1}{\rho_i-1}\cdot \frac1p\tr \left(\Sigma_i^{-1}\right)$ and $(\cal M_2)_{ii}=\frac{1}{\rho_i-1}\cdot \frac1p\tr \left(\Sigma_i^{-1}\Sigma_k\right)$.
\end{proposition}

\begin{proof}[Proof of Proposition \ref{prop_taskonomy}]
For each task $i$, $1\le i \le t-1$, we can find that  
$$\hat \beta_i = \beta_i + (X_i^\top X_i)^{-1}X_i^\top \e_i.$$
Then we calculate that
\begin{align*}
\|\hat\beta_i\|^2 &= \|\beta_i\|^2 + 2\beta_i^\top\cdot (X_i^\top X_i)^{-1}X_i^\top \e_i + \|(X_i^\top X_i)^{-1}X_i^\top \e_i \|^2 .
\end{align*}
We can use concentration of random vector with independent entries, Lemma \ref{largedeviation}, to get that for any deterministic vector $\beta$ with $\|\beta\|=\OO(1)$,
\be\label{largebetae}
\beta^\top (X_i^\top X_i)^{-1}X_i^\top \e_i  \le p^{\e/2}\cdot \sigma \left[\beta^\top (X_i^\top X_i)^{-1}\beta \right]^{1/2}\le \sigma p^{-1/2+\e},  
\ee
with high probability for any constant $\e>0$, where we used Lemma \ref{SxxSyy} to bound 
\be\label{opXiXi}\| (X_i^\top X_i)^{-1}\|\lesssim n_i^{-1} \quad \text{whp}. \ee
Similarly, using Lemma \ref{largedeviation} we get with high probability,
\be\label{largebetae2}
\begin{split}
 \left|\|(X_i^\top X_i)^{-1}X_i^\top \e_i \|^2 - \sigma^2 \tr\left( (X_i^\top X_i)^{-1} \right)\right|   \le p^{\e/2}\cdot \sigma^2 \left\{\tr\left[X_i (X_i^\top X_i)^{-2}X_i^\top\right]^2\right\}^{1/2} \\
 = p^{\e/2}\cdot \sigma^2 \left[ \tr (X_i^\top X_i)^{-2} \right]^{1/2} \le \sigma^2 p^{-1/2+\e},
 \end{split}
\ee 
and for $i\ne j$,
\be \label{largebetae3}
\begin{split}
\left| \e_j^\top X_j (X_j^\top X_j)^{-1}(X_i^\top X_i)^{-1}X_i^\top \e_i \right| &\le   p^{\e/2}\cdot \sigma^2 \left\{ \tr \left[(X_i^\top X_i)^{-1}(X_j^\top X_j)^{-1}\right] \right\}^{1/2} \\
&\le \sigma^2 p^{-1/2+\e},  
 \end{split}
\ee
%where in the we used Theorem \ref{LEM_SMALL} to get that
%$$\beta_t^\top (X_i^\top X_i)^{-1}\beta_t  \le p^{-1+\e/2} \quad \text{and}\quad \tr \left[ (X_i^\top X_i)^{-2} \right]  \le p^{-1+\e/2} \quad \text{whp}.$$
With \eqref{largebetae}-\eqref{largebetae3}, we get that with high probability,
\begin{align}
\|\hat\beta_i\|^2&= \|\beta_i\|^2 + \sigma^2\cdot \tr \left( (X_i^\top X_i)^{-1}\right) +\OO(\sigma p^{-1/2+\e}) \nonumber\\
&=  \|\beta_i\|^2 + \frac{\sigma^2}{\rho_i-1}\cdot \frac1p\tr \left(\Sigma_i^{-1}\right) +\OO(\sigma p^{-1/2+\e}) ,\label{betanormi}
\end{align}
and %where in the second step we used Lemma \ref{lem_minv}. Similarly, we have
\begin{align}\label{betanormij}
\hat\beta_i^\top \hat \beta_j &= \beta_i^\top \beta_j + \OO(\sigma p^{-1/2+\e})  , \quad i\ne j.
\end{align}
With \eqref{betanormi} and \eqref{betanormij}, we obtain that with high probability,
\be\label{BTB}
B^\top B= (B^\star_{t-1})^\top B^\star_{t-1} + \sigma^2\cal M_1 +\OO(\sigma p^{-1/2+\e}),
\ee
where $\OO(\sigma p^{-1/2+\e})$ means a $(t-1)\times (t-1)$ matrix, say $\cal E$, satisfying $\|\cal E\|\le C\sigma p^{-1/2+\e}$. Notice that by \eqref{BTBassm}, we also have with high probability,
$$\|(B^\top B)^{-1}\| \lesssim \left\|\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1}\right\|= \OO( 1).$$
Moreover, using \eqref{largebetae} we get that
\be\label{BTBeta}
B^\top \beta =  (B^\star_{t-1})^\top \beta + \OO(\sigma p^{-1/2+\e}),
\ee
for any deterministic vector $\beta$ with $\|\beta\|=\OO(1)$.

>>>>

\subsection{Proof of the Transfer Learning Setting}\label{app_proof_sec4}

\begin{proposition}\label{prop_taskonomy}
For $i=1,\cdot, t$, let $Y_i = X_i\beta_i + \varepsilon_i$ be independent data models, where $X_i=Z_i\Sigma_i^{1/2}\in \R^{n_i\times p}$ satisfy Assumption \ref{assm_secA1} with $\rho_i:=n_i/p>1$ being fixed constants, and $\e_i\in \R^{n_i}$ are  random vectors with i.i.d. entries with mean zero, variance $\sigma^2$ and all moments as in \eqref{assmAhigh}.  Moreover suppose that $X_i$, $\beta_i$ and $\varepsilon_i$ are independent of each other.
%Let $X_i \in\real^{n_i \times p}$ and $Y_i = X_i\beta_i + \varepsilon_i$, for $i = 1, \cdots, t$, where $\e_i$ are random vectors with i.i.d. entries of mean zero and variance $\sigma^2$. Let $\rho_i>1$, $1\le i \le t$ be fixed constants. 
Assume that $\max_{1\le i \le t}\|\beta_i\|=\OO(1)$, $\sigma^2=\OO(1)$ and 
\be\label{BTBassm}
\left\|\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1}\right\|=\OO(1),\quad B_{t-1}^{\star}: = [{\beta}_1,{\beta}_2,\dots,{\beta}_{t-1}].
\ee
Then we have with high probability,
\be\label{eq_taskonomy}
\te(B W_t) =\|\e_\beta + \sigma^2 \e_{\vari}^{(1)}\|^2+ \sigma^2 \|\e_{\vari}^{(2)}\|^2 + \OO(p^{-1/2+\e}),
\ee
where 
\begin{align*}
\e_\beta &:= \Sigma_t^{1/2}\left\{1 - B^\star_{t-1}\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1} (B^\star_{t-1})^\top\right\} \beta_t,\\
\e_{\vari}^{(1)}&:=  \Sigma_t^{1/2}  B^\star_{t-1}\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1}\cal M_1 \left[(B^\star_{t-1})^\top B^\star_{t-1} + \sigma^2\cal M_1\right]^{-1} (B^\star_{t-1})^\top \beta_t,\\
\e_{\vari}^{(2)}&:= \cal M_2^{1/2}\left[(B^\star_{t-1})^\top B^\star_{t-1} + \sigma^2\cal M_1\right]^{-1} (B^\star_{t-1})^\top \beta_t.
\end{align*}
Here $\cal M_1$ and $\cal M_2$ are $(t-1)\times (t-1)$ diagonal matrices with $(\cal M_1)_{ii}=\frac{1}{\rho_i-1}\cdot \frac1p\tr \left(\Sigma_i^{-1}\right)$ and $(\cal M_2)_{ii}=\frac{1}{\rho_i-1}\cdot \frac1p\tr \left(\Sigma_i^{-1}\Sigma_t\right)$.
\end{proposition}

\begin{proof}[Proof of Proposition \ref{prop_taskonomy}]
\todo{check!} For each task $i$, $1\le i \le t-1$, we can find that  
$$\hat \beta_i = \beta_i + (X_i^\top X_i)^{-1}X_i^\top \e_i.$$
Then we calculate that
\begin{align*}
\|\hat\beta_i\|^2 &= \|\beta_i\|^2 + 2\beta_i^\top\cdot (X_i^\top X_i)^{-1}X_i^\top \e_i + \|(X_i^\top X_i)^{-1}X_i^\top \e_i \|^2 .
\end{align*}
We can use concentration of random vector with independent entries, Lemma \ref{largedeviation}, to get that for any deterministic vector $\beta$ with $\|\beta\|=\OO(1)$,
\be\label{largebetae}
\beta^\top (X_i^\top X_i)^{-1}X_i^\top \e_i  \le p^{\e/2}\cdot \sigma \left[\beta^\top (X_i^\top X_i)^{-1}\beta \right]^{1/2}\le \sigma p^{-1/2+\e},  
\ee
with high probability for any constant $\e>0$, where we used Lemma \ref{SxxSyy} to bound 
\be\label{opXiXi}\| (X_i^\top X_i)^{-1}\|\lesssim n_i^{-1} \quad \text{whp}. \ee
Similarly, using Lemma \ref{largedeviation} we get with high probability,
\be\label{largebetae2}
\begin{split}
 \left|\|(X_i^\top X_i)^{-1}X_i^\top \e_i \|^2 - \sigma^2 \tr\left( (X_i^\top X_i)^{-1} \right)\right|   \le p^{\e/2}\cdot \sigma^2 \left\{\tr\left[X_i (X_i^\top X_i)^{-2}X_i^\top\right]^2\right\}^{1/2} \\
 = p^{\e/2}\cdot \sigma^2 \left[ \tr (X_i^\top X_i)^{-2} \right]^{1/2} \le \sigma^2 p^{-1/2+\e},
 \end{split}
\ee 
and for $i\ne j$,
\be \label{largebetae3}
\begin{split}
\left| \e_j^\top X_j (X_j^\top X_j)^{-1}(X_i^\top X_i)^{-1}X_i^\top \e_i \right| &\le   p^{\e/2}\cdot \sigma^2 \left\{ \tr \left[(X_i^\top X_i)^{-1}(X_j^\top X_j)^{-1}\right] \right\}^{1/2} \\
&\le \sigma^2 p^{-1/2+\e},  
 \end{split}
\ee
%where in the we used Theorem \ref{LEM_SMALL} to get that
%$$\beta_t^\top (X_i^\top X_i)^{-1}\beta_t  \le p^{-1+\e/2} \quad \text{and}\quad \tr \left[ (X_i^\top X_i)^{-2} \right]  \le p^{-1+\e/2} \quad \text{whp}.$$
With \eqref{largebetae}-\eqref{largebetae3}, we get that with high probability,
\begin{align}
\|\hat\beta_i\|^2&= \|\beta_i\|^2 + \sigma^2\cdot \tr \left( (X_i^\top X_i)^{-1}\right) +\OO(\sigma p^{-1/2+\e}) \nonumber\\
&=  \|\beta_i\|^2 + \frac{\sigma^2}{\rho_i-1}\cdot \frac1p\tr \left(\Sigma_i^{-1}\right) +\OO(\sigma p^{-1/2+\e}) ,\label{betanormi}
\end{align}
and %where in the second step we used Lemma \ref{lem_minv}. Similarly, we have
\begin{align}\label{betanormij}
\hat\beta_i^\top \hat \beta_j &= \beta_i^\top \beta_j + \OO(\sigma p^{-1/2+\e})  , \quad i\ne j.
\end{align}
With \eqref{betanormi} and \eqref{betanormij}, we obtain that with high probability,
\be\label{BTB}
B^\top B= (B^\star_{t-1})^\top B^\star_{t-1} + \sigma^2\cal M_1 +\OO(\sigma p^{-1/2+\e}),
\ee
where $\OO(\sigma p^{-1/2+\e})$ means a $(t-1)\times (t-1)$ matrix, say $\cal E$, satisfying $\|\cal E\|\le C\sigma p^{-1/2+\e}$. Notice that by \eqref{BTBassm}, we also have with high probability,
$$\|(B^\top B)^{-1}\| \lesssim \left\|\left[(B^\star_{t-1})^\top B^\star_{t-1}\right]^{-1}\right\|= \OO( 1).$$
Moreover, using \eqref{largebetae} we get that
\be\label{BTBeta}
B^\top \beta =  (B^\star_{t-1})^\top \beta + \OO(\sigma p^{-1/2+\e}),
\ee
for any deterministic vector $\beta$ with $\|\beta\|=\OO(1)$.
